---
title: "Getting and Cleaning Data"
author: "Konstantinos Papastamos"
date: "20 March 2015"
output: pdf_document
---

The script checks if the required dataset exists in your working directory.
If not it downloads the zip file, unzips it, deletes the zip file


```{r}
if(!file.exists("UCI HAR Dataset")){
  url="https://d396qusza40orc.cloudfront.net/"
  url = paste(url,"getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip",sep = "")
  download.file(url,"data")
  unzip("data")
  file.remove("data")
}

```

```{r "setup", include=FALSE}
knitr::opts_knit$set(root.dir = "UCI HAR Dataset")
```

Checks if package plyr is installed(we are going to need it later) and if not installs it and 
loads it, if yes it loads it.

```{r}
if(require(plyr)==FALSE){
  install.packages("plyr")
}else{
  library(plyr)
}
```


Reads the activity ,the feature and the subject files and stores
them in the corresponding variables


```{r}
train_activities=read.table("train\\y_train.txt",header=FALSE)
test_activities=read.table("test\\y_test.txt",header=FALSE)



train_features=read.table("train\\X_train.txt",header=FALSE)
test_features=read.table("test\\X_test.txt",header=FALSE)



train_subject=read.table("train\\subject_train.txt",header=FALSE)
test_subject=read.table("test\\subject_test.txt",header=FALSE)
```

Merges the two datasets and names the variables accordingly


```{r}
subject=rbind(train_subject,test_subject)
features=rbind(train_features,test_features)
activities=rbind(train_activities,test_activities)

names(subject)="subjects"
names(activities)="activities"
feature_names=read.table("features.txt")
feature_names=feature_names[,2]
names(features)=feature_names

```


Creates the final dataset by adding the subject,features and activities dataframes together

```{r}
full_dataset=cbind(subject,activities,features)

```

Reads the activity labels from the activity_labels.txt file and changes activity labels class
in the dataset to factor. Then changes the factor's levels from 1,2,3,4,5,6 to the appropriate
label

```{r}
activity_labels=read.table("activity_labels.txt",header=FALSE)

full_dataset[,2]=as.factor(full_dataset[,2])

levels(full_dataset[,2])=activity_labels$V2

```


Labels the data with descriptive variable names, changing the t to time, the f to frequency, the
Acc to Accelerometer, the Gyro to Gyroscope, the BodyBody to Body,the Mag to Magnitude and the
tBody to timeBody(some variables did't begin with the letter t so i needed to add this last one)

```{r}
names(full_dataset)=gsub("^f","frequency",names(full_dataset))
names(full_dataset)=gsub("^t","time",names(full_dataset))
names(full_dataset)=gsub("Acc","Accelerometer",names(full_dataset))
names(full_dataset)=gsub("Gyro","Gyroscope",names(full_dataset))
names(full_dataset)=gsub("BodyBody","Body",names(full_dataset))
names(full_dataset)=gsub("Mag","Magnitude",names(full_dataset))
names(full_dataset)=gsub("tBody","timeBody",names(full_dataset))
```

Creates a second, independent tidy data set with the average of each variable for each activity 
and each subject

```{r}
tidy_dataset=aggregate(. ~ subjects+activities,full_dataset,mean)
tidy_dataset=arrange(tidy_dataset,subjects,activities)
```
